Changelog

### 2021-05-19 Chapter-2/Chapter-2.tex
Updates from Overleaf

Based on the above attacks, some work~\cite{brickell2011technologies} suggests having secret-dependent memory accesses at finer than the size of one cache line should be secure enough. However, recent studies show that having different memory accesses within a cache line is also dangerous. One example is the CacheBleed attack~\cite{yarom2017cachebleed}. It is a cache channel attack that can exploit cache-bank collisions. 
\subsection{Page-level Channel}
Page channel attacks allow an attacker can observe the memory access at the granularity of memory pages. One example is the controlled-channel attack~\cite{xu2015controlled}. It is a primary threat to the trusted execution environments (TEEs).  A privileged adversary (e.g., kernel) can infer sensitive information by manipulating the enclave page accesses. A malicious OS can proactively revoke a virtual enclave memory page from the virtual memory. If the victim enclave program accesses the memory in the  page, then the OS can observe the page faults and know which page is accessed. While the granularity of the controlled-channel attacks is more coarse-grained compared with cache channel attacks, controlled-channel attacks allow the attacker to have a noise-free observation. Other examples include the TLB side-channel attack and DDRAM attacks share the same page-level granularity.

@@

\subsection{Attack Models}
Previous work~\cite{182946} classifies side-channel attacks into three types based on the type of information that an attacker can learn during the attack.
\subsubsection*{Timing-based Attacks}
In timing-based attacks, an adversary can monitor the execution time of a victim program. It could be the overall execution time or the execution time of some interesting points (e.g., a function).  The execution time is correlated to some events (e.g., cache hits and misses, page faults, the number of instructions). Examples of those attacks include remote timing attacks~\cite{brumley2005remote}, Bernstein's attack on AES~\cite{bernstein2005cache}, and Kocher’s original attack~\cite{kocher1996timing} from CRYPTO 96.
\subsubsection*{Access-based Attacks}
In access-based attacks, an adversary can probe one time only after the termination of the victim program. The attacker can learn the memory addresses that are accessed during the execution. For example, a cache-based channel attack can know which memory access incurs a cache hit and which memory access incurs a cache miss. Since the attacker can probe the victim program after the execution, the attacker does not know the order of memory accesses. Access-based attacks often happen in the cloud computing environment when the attacker and victim share the same hardware (e.g., LLC cache). Examples of those attacks include cache template attacks~\cite{191010}, Prime and Probe attacks~\cite{liu2015last}, and Flush and Reload attacks~\cite{yarom2014flush+}.
\subsubsection*{Trace-based Attacks}
In trace-based attacks, an adversary can probe the victim program at any point. The attacker can learn the order of memory accesses. Suppose the victim program has two functions \textsf{a} and \textsf{b}. In an access-based attacks, the attacker can only learn both \textsf{a} and \textsf{b} are executed. In a trace-based attack, the attacker can know the execution sequence of two functions (e.g., $aabbb$). Controlled-channel attacks~\cite{moghimi2020copycat, xu2015controlled} belong to the category where a malicious operating system can interrupt the victim application and probe the program at any time. In general, the trace-based attack has the strongest assumption about the threat model,  but it can retrieve more information compared to the other two types of attacks.
\subsection{Target Applications}
There is a vast number of studies\cite{yarom2017cachebleed,191010,184415,Osvik2006,liu2015last} on side-channel attacks. Most early work targeted retrieving encryption keys from cryptography libraries. In recent years, researchers start to focus on more general and diverse applications. In summary,  the target applications of current side-channel attacks can be classified into three categories.

@@

\section{Trace-based Binary Analysis}
Trace-based binary analysis can be seen as a type of dynamic analysis. It analyzes one concrete program path one time rather than trying to cover
as many paths as possible. Compared to static analysis, trace-based analysis
is usually more precise as it can incorporate runtime information. It usually has
two steps. The first step is collecting trace information. Existing
approaches usually often rely on binary instrumentation frameworks or emulators (e.g., Intel Pin~\cite{luk2005pin},
Valgrind~\cite{nethercote2007valgrind}, QEMU). The second step is analyzing the trace using techniques, such
as symbolic execution, taint analysis. Depending on when the analysis is launched,
there are two kinds of trace-based analysis, online analysis, and offline analysis. As the name suggests, online analysis is performed during the program execution. The target program's data can be modified directly in memory, which is useful to tasks like fuzzing because the analysis
can force the program to execute a specific branch. 
Offline analysis or post-analysis happens after the program finishes executions. 
Trace-based binary analysis has been adopted in several tasks like algorithm
detection, vulnerability detection, and malware analysis. We only mention a few of
the most related works here. Xu et al.~\cite{xu2017cryptographic} introduce bitwise symbolic execution, which can
identify crypto algorithms in obfuscated binaries. Ming et al.~\cite{ming2017binsim} use system call sliced segment equivalence checking to do binary diffing. Jia et al.~\cite{jia2017towards}
perform an offline analysis on execution traces to identify heap overflows.

>>>>

Based on the above attacks, some work~\cite{brickell2011technologies} suggests having secret-dependent memory accesses at finer than the size of one cache line should be secure enough. However, recent studies show that having different memory accesses within a cache line is also dangerous. One example is the CacheBleed attack~\cite{yarom2017cachebleed}. It is a cache channel attack that can exploit cache-bank collisions. 
\subsection{Page-level Channel}
Page channel attacks allow an attacker can observe the memory access at the granularity of memory pages. One example is the controlled-channel attack~\cite{xu2015controlled}. It is a primary threat to the trusted execution environments (TEEs).  A privileged adversary (e.g., kernel) can infer sensitive information by manipulating the enclave page accesses. A malicious OS can proactively revoke a virtual enclave memory page from the virtual memory. If the victim enclave program accesses the memory in the  page, then the OS can observe the page faults and know which page is accessed. While the granularity of the controlled-channel attacks is more coarse-grained compared with cache channel attacks, controlled-channel attacks allow the attacker to have a noise-free observation. Other examples such as TLB side-channel attacks and DDRAM attacks share the same page-level granularity.

@@

\subsection{Attack Models}
Previous work~\cite{182946} classifies side-channel attacks into three types based on the type of information that an attacker can learn during the attack.
\subsubsection*{Timing-based Attacks}
In timing-based attacks, an adversary can monitor the execution time of a victim program. It could be the overall execution time or the execution time of some interesting points (e.g., a function).  The execution time is correlated to some events (e.g., cache hits and misses, page faults, the number of instructions). Examples of those attacks include remote timing attacks~\cite{brumley2005remote}, Bernstein's attack on AES~\cite{bernstein2005cache}, and Kocher’s original attack~\cite{kocher1996timing} from CRYPTO 96.
\subsubsection*{Access-based Attacks}
In access-based attacks, an adversary can probe one time only after the termination of the victim program. The attacker can learn the memory addresses that are accessed during the execution. For example, a cache-based channel attack can know which memory access incurs a cache hit and which memory access incurs a cache miss. Since the attacker can probe the victim program after the execution, the attacker does not know the order of memory accesses. Access-based attacks often happen in the cloud computing environment when the attacker and victim share the same hardware (e.g., LLC cache). Examples of those attacks include cache template attacks~\cite{191010}, Prime+Probe attacks~\cite{liu2015last}, and Flush+Reload attacks~\cite{yarom2014flush+}.
\subsubsection*{Trace-based Attacks}
In trace-based attacks, an adversary can probe the victim program at any point. The attacker can learn the order of memory accesses. Suppose the victim program has two functions \textsf{a} and \textsf{b}. In an access-based attacks, the attacker can only learn both \textsf{a} and \textsf{b} are executed. In a trace-based attack, the attacker can know the execution sequence of two functions (e.g., $aabbb$). Controlled-channel attacks~\cite{moghimi2020copycat, xu2015controlled} belong to the category where a malicious operating system can interrupt the victim application and probe the program at any time. In general, the trace-based attack has the strongest assumption about the threat model,  but it can retrieve more information compared to the other two types of attacks.
\subsection{Target Applications}
There is a vast number of studies\cite{yarom2017cachebleed,191010,184415,Osvik2006,liu2015last} on side-channel attacks. Most early work targeted retrieving encryption keys from cryptography libraries. In recent years, researchers start to focus on more general and diverse applications. In summary,  the target applications of current side-channel attacks can be classified into three categories.

@@

\section{Trace-based Binary Analysis}
Trace-based binary analysis can be seen as a type of dynamic analysis. It analyzes one concrete program path one time rather than trying to cover
as many paths as possible. Compared to static analysis, trace-based analysis
is usually more precise as it can incorporate runtime information. It usually has
two steps. The first step is collecting trace information. Existing
approaches usually often rely on binary instrumentation frameworks or emulators (e.g., Intel Pin~\cite{luk2005pin},
Valgrind~\cite{nethercote2007valgrind}, and QEMU). The second step is analyzing the trace using techniques, such
as symbolic execution, taint analysis. Depending on when the analysis is launched,
there are two kinds of trace-based analysis, online analysis, and offline analysis. As the name suggests, online analysis is performed during the program execution. The target program's data can be modified directly in memory, which is useful to tasks like fuzzing because the analysis
can force the program to execute a specific branch. 
Offline analysis or post-analysis happens after the program finishes executions. 
Trace-based binary analysis has been adopted in several tasks like algorithm
detection, vulnerability detection, and malware analysis. We only mention a few of
the most related works here. Xu et al.~\cite{xu2017cryptographic} introduce bitwise symbolic execution, which can
identify crypto algorithms in obfuscated binaries. Ming et al.~\cite{ming2017binsim} use system call sliced segment equivalence checking to do binary diffing. Jia et al.~\cite{jia2017towards}
perform an offline analysis on execution traces to identify heap overflows.
